Applying LIME to Bullet Data Journal

### 2018-12-12 journal/applying LIME to bullet data journal.Rmd
finished journal entry for this week

# December 5, 2018 to December 11, 2018

@@

```{r}
hamby224_lime_results <- hamby224_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff^2)) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(set) %>%
  ungroup() %>%
  mutate(nbins = as.numeric(as.character(nbins))) %>%
  mutate(situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE, 
                            sprintf("%.0f quantile bins", nbins),
                            ifelse(bin_continuous == TRUE & quantile_bins == FALSE, 
                                   sprintf("%.0f equally spaced bins", nbins),
                                   ifelse(bin_continuous == FALSE & use_density == TRUE, 
                                          "kernel density", "normal approximation")))) %>%
  mutate(situation = fct_relevel(situation, 
                                 "2 quantile bins", 
                                 "3 quantile bins",
                                 "4 quantile bins",
                                 "5 quantile bins",
                                 "6 quantile bins"))
```

@@

```{r}
hamby224_lime_results %>%
  ggplot(aes(x = situation, y = mse)) +
  geom_point() + 
  facet_grid( ~ set) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r}
hamby224_lime_results %>%
  ggplot(aes(x = situation, y = ave_r2)) +
  geom_point() + 
  facet_grid( ~ set) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r}
hamby224_sensitivity_results <- hamby224_sensitivity_outputs %>%
  mutate(situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE, 
                            sprintf("%.0f quantile bins", nbins),
                            ifelse(bin_continuous == TRUE & quantile_bins == FALSE, 
                                   sprintf("%.0f equally spaced bins", nbins),
                                   ifelse(bin_continuous == FALSE & use_density == TRUE, 
                                          "kernel density", "normal approximation"))),
         diff = rfscore - model_prediction) %>%
  mutate(situation = fct_relevel(situation, 
                                 "2 quantile bins", 
                                 "3 quantile bins",
                                 "4 quantile bins",
                                 "5 quantile bins",
                                 "6 quantile bins"))
```

```{r}
hamby224_sensitivity_mses <- hamby224_sensitivity_results %>%
  group_by(set, rep, situation) %>%
  summarise(mse = sum((diff)^2) / length(diff))
```

@@

```{r}
hamby224_sensitivity_mses %>%
  ggplot(aes(x = situation, y = mse)) + 
  geom_boxplot() + 
  geom_point() + 
  facet_grid( ~ set) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "", y = "MSE")
```

```{r}
hamby224_consistency <- hamby224_sensitivity_results %>%
  select(set, situation, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(set, situation, case, rep, desc(feature_weight_abs)) %>%
  group_by(set, situation, case, rep) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(set, situation, case) %>%
  summarise(nlevels = length(levels(factor(feature))),
            count = n()) %>%
  ungroup() %>%
  group_by(set, situation) %>%
  summarise(mean_nlevels = mean(nlevels),
            sd_nlevels = sd(nlevels),
            upper = mean(nlevels) + sd(nlevels),
            lower = mean(nlevels) - sd(nlevels),
            max = max(nlevels),
            min = min(nlevels))
```

@@

```{r}
hamby224_consistency %>%
  ggplot(aes(x = situation, y = mean_nlevels)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  facet_grid(. ~ set) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r}
hamby224_sensitivity_results %>%
  select(set, situation, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(set, situation, case, rep, desc(feature_weight_abs)) %>%
  group_by(set, situation, case, rep) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(set, situation, case) %>%
  summarise(nlevels = length(levels(factor(feature)))) %>%
  ggplot(aes(x = situation, fill = factor(nlevels))) + 
  geom_bar(position = "stack") +
  facet_grid( ~ set) + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

@@

@@

```{r}
joined_results <- hamby224_lime_results %>%
  ungroup() %>%
  mutate(nbins = as.numeric(as.character(nbins))) %>%
  mutate(situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE, 
                            sprintf("%.0f quantile bins", nbins),
                            ifelse(bin_continuous == TRUE & quantile_bins == FALSE, 
                                   sprintf("%.0f equally spaced bins", nbins),
                                   ifelse(bin_continuous == FALSE & use_density == TRUE, 
                                          "kernel density", "normal approximation")))) %>%
  mutate(situation = fct_relevel(situation, 
                                 "2 quantile bins", 
                                 "3 quantile bins",
                                 "4 quantile bins",
                                 "5 quantile bins",
                                 "6 quantile bins")) %>%
  select(set, situation, mse, ave_r2) %>%
  full_join(hamby224_sensitivity_mses %>%
              group_by(set, situation) %>%
              summarise(mean_sensitivity_mse = mean(mse), 
                        sd_sensitivity_mse = sd(mse)),
            by = c("set", "situation")) %>%
  full_join(hamby224_consistency %>% select(set, situation, mean_nlevels, sd_nlevels), 
            by = c("set", "situation")) %>%
  #filter(!(situation %in% c("normal approximation", "kernel density"))) %>%
  group_by(set) %>%
  mutate(mse_order = dense_rank(mse),
         ave_r2_order = dense_rank(desc(ave_r2)),
         mean_sensitivity_mse_order = dense_rank(mean_sensitivity_mse),
         mean_nlevels_order = dense_rank(mean_nlevels))
```

```{r}
joined_results %>%
  ungroup() %>%
  filter(set == "Set 1") %>%
  select(situation, mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  arrange(mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  knitr::kable()

joined_results %>%
  ungroup() %>%
  filter(set == "Set 11") %>%
  select(situation, mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  arrange(mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  knitr::kable()
```

>>>>

# December 5, 2018 to December 14, 2018

@@

```{r}
hamby224_lime_results <- hamby224_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff^2)) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(set) %>%
  ungroup() %>%
  mutate(nbins = as.numeric(as.character(nbins))) %>%
  mutate(situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE, 
                            sprintf("%.0f quantile bins", nbins),
                            ifelse(bin_continuous == TRUE & quantile_bins == FALSE, 
                                   sprintf("%.0f equally spaced bins", nbins),
                                   ifelse(bin_continuous == FALSE & use_density == TRUE, 
                                          "kernel density", "normal approximation")))) %>%
  mutate(situation = fct_relevel(situation, 
                                 "2 quantile bins", 
                                 "3 quantile bins",
                                 "4 quantile bins",
                                 "5 quantile bins",
                                 "6 quantile bins")) %>%
  mutate(bin_situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE,
                                "quantile bins",
                                ifelse(bin_continuous == TRUE & quantile_bins == FALSE,
                                       "equally spaced bins", "other")))
```

@@

```{r}
hamby224_lime_results %>%
  ggplot(aes(x = situation, y = mse, color = bin_situation)) +
  geom_point() + 
  facet_grid( ~ set) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "bottom")
```

@@

```{r}
hamby224_lime_results %>%
  ggplot(aes(x = situation, y = ave_r2, color = bin_situation)) +
  geom_point() + 
  facet_grid( ~ set) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

@@

```{r}
hamby224_sensitivity_results <- hamby224_sensitivity_outputs %>%
  mutate(situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE, 
                            sprintf("%.0f quantile bins", nbins),
                            ifelse(bin_continuous == TRUE & quantile_bins == FALSE, 
                                   sprintf("%.0f equally spaced bins", nbins),
                                   ifelse(bin_continuous == FALSE & use_density == TRUE, 
                                          "kernel density", "normal approximation"))),
         diff = rfscore - model_prediction) %>%
  mutate(situation = fct_relevel(situation, 
                                 "2 quantile bins", 
                                 "3 quantile bins",
                                 "4 quantile bins",
                                 "5 quantile bins",
                                 "6 quantile bins")) %>%
  mutate(bin_situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE,
                                "quantile bins",
                                ifelse(bin_continuous == TRUE & quantile_bins == FALSE,
                                       "equally spaced bins", "other")))
```

```{r}
hamby224_sensitivity_mses <- hamby224_sensitivity_results %>%
  group_by(set, rep, situation, bin_situation) %>%
  summarise(mse = sum((diff)^2) / length(diff))
```

@@

```{r}
hamby224_sensitivity_mses %>%
  ggplot(aes(x = situation, y = mse)) + 
  geom_boxplot(aes(color = bin_situation)) + 
  geom_point() + 
  facet_grid( ~ set) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "", y = "MSE")
```

```{r}
hamby224_consistency <- hamby224_sensitivity_results %>%
  select(set, situation, bin_situation, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(set, situation, case, rep, desc(feature_weight_abs)) %>%
  group_by(set, situation, case, rep) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(set, situation, bin_situation, case) %>%
  summarise(nlevels = length(levels(factor(feature))),
            count = n()) %>%
  ungroup() %>%
  group_by(set, situation, bin_situation) %>%
  summarise(mean_nlevels = mean(nlevels),
            sd_nlevels = sd(nlevels),
            upper = mean(nlevels) + sd(nlevels),
            lower = mean(nlevels) - sd(nlevels),
            max = max(nlevels),
            min = min(nlevels))
```

@@

```{r}
hamby224_consistency %>%
  ggplot(aes(x = situation, y = mean_nlevels)) + 
  geom_point(aes(color = bin_situation)) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  facet_grid(. ~ set) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r}
hamby224_sensitivity_results %>%
  select(set, situation, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(set, situation, case, rep, desc(feature_weight_abs)) %>%
  group_by(set, situation, case, rep) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(set, situation, case) %>%
  summarise(nlevels = length(levels(factor(feature)))) %>%
  ggplot(aes(x = situation, fill = factor(nlevels))) + 
  geom_bar(position = "stack") +
  facet_grid( ~ set) + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_brewer(palette = "Blues")
```

@@

The fact that different features get chosen as the top feature for different estimation techniques makes sense. You would expect that the number of bins would determine which features are better suited for predicting whether or not a comparison is a match. For example, if you look back at the feature distribution plots, you can see that ccf is the obvious choice for determining between matches and non-matches if you have to choose two equally spaced bins. This is the feature that lime most frequently chooses. You can make similar arguments for other cases as well. Even though different inputs select different variables, I would expect that some variables will be better at prediction than others. I think this is what leads some input values to better LIME explanations and lower MSEs than others.

@@

I was thinking a bit more about the relationship between the top features chosen and the type of binning used. I realized it would be helpful to view the relationship between the features and the random forest score. The plots below show the testing data with random forest score versus each of the features. The points are colored by the samesource variable.

```{r}
hamby224_test %>% 
  select(rf_features, rfscore, samesource) %>%
  gather(key = feature, value = value, 1:9) %>%
  select(feature, value, rfscore, samesource) %>%
  ggplot(aes(x = value, y = rfscore, color = samesource)) + 
  geom_point() + 
  facet_wrap( ~ feature, scales = "free") +
  labs(x = "Variable Value", y = "Random Forest Score", color = "Same Source?",
       title = "Hamby 224 Testing Data") +
  theme_bw() +
  theme(legend.position = "bottom")

```

@@

While using these to help me decide which input settings to use, I had a handful of thoughts:

* Since I am still working on making sense of the variation results, and it would be nice to get the first paper out, I am thinking of basing the decision of input values entirely off of only the assessment measures from the one case.
* Even with needing more time to think about how to understand the meaning of the results from the 10 reps, the mean MSE from these has very similar results to the MSEs from the one rep, and it does not appear that the top chosen features vary too much between the reps. This makes me think that it would be okay to base my decision off of the results from the single rep.
* Both Heike and I do not think that the $R^2$ is that helpful of a measure for determining whether the LIME explanations are good, so I am not going to put much weight on it.
* Additionally, I do not want to consider either the normal approximation or kernel density estimate since I have not figured out a good way to visualize these in the app. Right now, the feature plots from these explanations are not very helpful. I would need to adjust the intercept or something.

Based on the MSEs from the single impelementation of LIME, I will use 3 equally spaced bins for both the set 1 and set 11 explanations in the firearm examiner paper since it leads to the lowest MSE for both sets.

```{r}
joined_results <- hamby224_lime_results %>%
  ungroup() %>%
  mutate(nbins = as.numeric(as.character(nbins))) %>%
  mutate(situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE, 
                            sprintf("%.0f quantile bins", nbins),
                            ifelse(bin_continuous == TRUE & quantile_bins == FALSE, 
                                   sprintf("%.0f equally spaced bins", nbins),
                                   ifelse(bin_continuous == FALSE & use_density == TRUE, 
                                          "kernel density", "normal approximation")))) %>%
  mutate(situation = fct_relevel(situation, 
                                 "2 quantile bins", 
                                 "3 quantile bins",
                                 "4 quantile bins",
                                 "5 quantile bins",
                                 "6 quantile bins")) %>%
  select(set, situation, mse, ave_r2) %>%
  full_join(hamby224_sensitivity_mses %>%
              group_by(set, situation) %>%
              summarise(mean_sensitivity_mse = mean(mse), 
                        sd_sensitivity_mse = sd(mse)),
            by = c("set", "situation")) %>%
  full_join(hamby224_consistency %>% select(set, situation, mean_nlevels, sd_nlevels), 
            by = c("set", "situation")) %>%
  group_by(set) %>%
  mutate(mse_order = dense_rank(mse),
         ave_r2_order = dense_rank(desc(ave_r2)),
         mean_sensitivity_mse_order = dense_rank(mean_sensitivity_mse),
         mean_nlevels_order = dense_rank(mean_nlevels))
```

```{r}
joined_results %>%
  ungroup() %>%
  filter(set == "Set 1") %>%
  select(situation, mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  arrange(mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  DT::datatable(options = list(pageLength = 12, paging = FALSE, searching = FALSE),
                rownames = FALSE,
                colnames = c("Situation", "MSE", "Mean MSE", "Mean Number of Top Features", 
                             "Mean R2"),
                caption = "Results from Set 1")

joined_results %>%
  ungroup() %>%
  filter(set == "Set 11") %>%
  select(situation, mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  arrange(mse_order, mean_sensitivity_mse_order, mean_nlevels_order, ave_r2_order) %>%
  DT::datatable(options = list(pageLength = 12, paging = FALSE, searching = FALSE),
                rownames = FALSE,
                colnames = c("Situation", "MSE", "Mean MSE", "Mean Number of Top Features", 
                             "Mean R2"),
                caption = "Results from Set 11")
```
